---
title: `PM2.5_attr_mort` is now an R package!
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{welcome}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Gener
For several years, the project has been encapsulated in the form of Rproject, requiring the entire project folder to be cloned each time when used to ensure that the data format and output locations remain consistent. 

This has caused significant inconvenience in flexibly applying the project's underlying methodology for calculating attributable mortality.

In the latest pushed version, I have refactored the core code that this project relies on, allowing it to take advantage of the latest capabilities of the tidyverse packages. It has become more object-oriented and no longer strictly adheres to the initial design of grid-based computation, making it suitable for more flexible data resolutions.

Moreover, installing a package is much more flexible than cloning the entire project—offering more adaptable function calls, more versatile update methods, and more flexible data imports.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Installation
I've yet not considering putting this project into CRAN until it gets more well-functioned, so at present, it can only installed through github.
```
devtools::install_github("evanliu3594/PM2.5-attr-mort")    
library(AttrMort)
```

# Data importing
As I mentioned earlier, the functions in the `AttrMort` package accept a more flexible arrangement of data than before. 
There is no longer a strict requirement for the input data to maintain a grid format of x-y. As long as the field data, population data, dose data, age structure data, and mortality data can be matched and combined according to some key columns, the calculations can be completed.
This package includes some sample data. Please refer to the following for more information.
```{r}
library(stringr)
builtin_datas <- list.files(system.file('extdata', package = "AttrMort"), full.names = T)

for (f in builtin_datas) {
  nm <- basename(f) %>% str_extract(".+(?=\\.)")
  assign(nm, readRDS(f))
}

head(Grid_info)
```

# Calculating Attributable Mortality
```{r}
library(AttrMort)

set_Model("NCD+LRI")

Mortality(
  field = Grid_info %>% filter(location == "China"),
  dose_real = get_at(Grid_PM25, scenario, "SSP1-Baseline_2030"),
  pop = get_at(Grid_Pop, scenario, "SSP1-Baseline_2030"),
  age = get_at(GBD_agestructure, scenario, "SSP1-Baseline_2030"),
  mort = get_at(GBD_mortrate, scenario, "SSP1-Baseline_2030") %>% mutate(endpoint = tolower(endpoint))
)
```

# Calculating Aggregated Attributable Mortality with uncertainty

```{r}
Mortality_Aggr(
  field = Grid_info %>% filter(location == "China"),
  dose_real = get_at(Grid_PM25, scenario, "base2015"),
  pop = get_at(Grid_Pop, scenario, "base2015"),
  age = get_at(GBD_agestructure, scenario, "base2015"),
  mort = get_at(GBD_mortrate, scenario, "base2015") %>% mutate(endpoint = tolower(endpoint)),
  doseRSME = 12, 
  lvl = "location", 
  uncertain = T
)
```

